---
title: "Open Science: git"
author: "Helen Kettle"
date: "07/04/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is git? 

Git is version control software i.e. it can be used as a way to back
up your work, collaborate with others and to go back to earlier
versions if you need to. The files are kept in repositories ("repos"
to the cool kids!) - the local repo is in your local file storage and
the web-based repo (the 'remote') is hosted on a server, for example
<https://github.com>. The repo on GitHub can be open to the public or
kept private. You may have used GitHub in the past as somewhere to put
your code/data for journal publications but you may not have used it
when developing your work. This page gives some links to more info on
git and a brief overview of how to use it once you have git on your
machine.

### Some links

<https://git-scm.com/book/en/v2>

<https://swcarpentry.github.io/git-novice/>

<https://www.freecodecamp.org/news/git-and-github-for-beginners/>

<https://www.muscardinus.be/2017/08/git_workflow_ecologists/>


### How to use git

I will give a brief overview below but see
<https://git-scm.com/book/en/v2> for more details.

Firstly make sure you have git on your machine (try typing 'git') and
then declare a folder/directory on your machine as a git repo (this is
done locally - no connection with the internet required!). To do this
on linux, go into the folder you want and type:

```{r eval=FALSE}
git init
```

To save files to your local repo you need to tell git to add the files
and then you do a commit (you must add a comment describing this
commit). So basically it is a two stage process to save your files to
the repo each time, eg. to add the Rscripts in your folder:

```{r eval=FALSE}
git add *.R

git commit -m 'add all my Rscripts'
```

You can view what git believes has changed at any point with:

```{r eval=FALSE}
git status
```

If you want to then copy these files to the web based repo on github,
you go to <https://github.com> and make a new repository. 

Then in your local repository create a remote branch with the short
name 'origin' (for example). 

Finally, 'push' your files to it.

```{r eval=FALSE}
git remote add origin https://github.com/HelenKettle/OpenScience.git 

git push origin master 
```

Note, you will need a personal access token (PAT) to push to the web
repo - see
<https://docs.github.com/en/enterprise-server@3.4/authentication/keeping-your-account-and-data-secure/creating-a-personal-access-token>


From now on, you can work as usual and when you want to commit to the
local repo you simply add the appropriate files, commit them, then
push to the remote repo as required:

```{r eval=FALSE}
git add *.R

git commit -m 'some new things happened' ##do not use this comment!

git push origin master 

```

